Approximation as Differentiation: Notes

### 2025-05-16 main/formal-structures.tex
LatGal has biproducts.

One observation about the (bounded) meet-semilattice homomorphisms from $X$ to $Y$ is that they admit a
meet-semilattice structure themselves: the pointwise lifting of $\meet$ to meet-preserving maps $f, g: X \to
Y$, where $(f \meet g)(x) = f(x) \meet g(x): X \to Y$ also preserves meets, giving a notion of $\meet$ for
morphisms, as does the constant map $\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$, and which acts a
unit for $\meet$. Dually the (bounded) join-preserving maps from $Y$ to $X$ also have a join-semilattice
structure given pointwise by $\join$ and the constant homomorphism $\bot_{Y,X}: Y \to X$ sending any $y \in Y$
to $\bot_Y$. Since these two constructions are adjoint, $\LatGal$ is enriched in the category $\SemiLat$ of
semilattices and semilattice homomorphisms. The hom-set $\LatGal(X,Y)$ forms a bounded meet semilattice with
unit $\top_{X,Y}$ given by the Galois connection $\bot_{Y,X} \adj \top_{X,Y}: X \to Y$ and meet of Galois
connections $(f \adj g) \meet (f' \adj g') = (f \join f') \adj (g \meet g'): X \to Y$.

Another notable property of $\LatGal$ is that the projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y
\to Y$, where $X \times Y$ denotes the product of lattices, have both upper and lower adjoints. This means $X
\times Y$, which we shall hereafter write as $X \biprod Y$, acts as both a product and a coproduct, with
projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$.

@@

\noindent The trivial 1-point lattice, which is both terminal and initial, is the unit for $\biprod$. Although
the presence of biproducts is incompatible with $\LatGal$ itself being Cartesian closed, the biproduct
structure will prove crucial for a establishing a Cartesian closed setting for interpreting higher-order
programs for Galois slicing.

>>>>

One observation about the (bounded) meet-semilattice homomorphisms from $X$ to $Y$ is that they admit a
meet-semilattice structure themselves: the pointwise lifting of $\meet$ to meet-preserving maps $f, g: X \to
Y$, where $(f \meet g)(x) = f(x) \meet g(x): X \to Y$ also preserves meets, giving a notion of $\meet$ for
morphisms, as does the constant map $\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$, and which acts a
unit for $\meet$. Dually the (bounded) join-preserving maps from $Y$ to $X$ have a join-semilattice structure
given pointwise by $\join$ and the constant homomorphism $\bot_{Y,X}: Y \to X$ sending any $y \in Y$ to
$\bot_X$. Since these two constructions are adjoint, $\LatGal$ is enriched in the category $\SemiLat$ of
semilattices and semilattice homomorphisms. The hom-set $\LatGal(X,Y)$ forms a bounded meet semilattice with
unit $\top_{X,Y}$ given by the Galois connection $\bot_{Y,X} \adj \top_{X,Y}: X \to Y$ and meet of Galois
connections $(f \adj g) \meet (f' \adj g') = (f \join f') \adj (g \meet g'): X \to Y$.

Another notable property of $\LatGal$ is that the projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y
\to Y$, where $X \times Y$ denotes the product of lattices, have both upper and lower adjoints. This means $X
\times Y$, which we shall hereafter write as $X \biprod Y$, acts as both a product and a coproduct, with
projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$ given by:

@@

\noindent The trivial 1-point lattice, which is both terminal and initial, is the unit for $\biprod$. Although
the presence of biproducts is incompatible with $\LatGal$ itself being Cartesian closed, the biproduct
structure will prove crucial for establishing a Cartesian closed setting for interpreting higher-order
programs for Galois slicing.
